---
title: "Konigsberg et al. (1998) stature"
author: "Lyle W. Konigsberg"
date: "`r Sys.Date()`"
bibliography: bibliography.bibtex
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Konigsberg et al. (1998) stature}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=T}
library(CRuff)
```

We will first do the default call to "K_run" which uses the femur equation from @RN10592 and applies it to the "European" data set from @RN12182. See "European data set: Excel" and "Notes" under "Special Projects" [here](https://fae.johnshopkins.edu/chris-ruff/).  Now for a quick word about @RN10592. There are not actually any such equations in @RN10592. Instead, the classical calibration method given in @RN8707 is used to calculate equations from the summary multivariate statistics in @RN10592.

```{r fig.dim=c(6,6)}
K_run()
```

Let's see for the total data set whether the proportion of cases in the 50 percent prediction interval actually contains 50 percent of the data.  That is a straight-forward binomial test:
```{r }
sto=156+636+270
print(sto)
print(binom.test(636,sto,0.5))
```
OK, not good.  The probability value is quite low, because the 95% confidence interval for the proportion in the 50% prediction interval is from about 0.5687 to 0.6285. It is even worse if we compare the proportion of cases below the 50% PIs to the total of those above and those below.
```{r }
sto=156+270
print(sto)
print(binom.test(156,sto,0.5))
```
OK, so that shows that the bulk of the cases (actual statures) are below the 50% PIs.  We would have expected about half should be below and half should be above.  But, that is not happening because the estimated statures are biased.  Generally, the estimates are greater than the actual statures, as in the plot.

Now let's use the female "short" data and the humerus.

```{r fig.dim=c(6,6)}
K_run(i.which="Hum",which.dat=2)
```

Now let's do the binomial test for the proportion in the 50% prediction intervals for the overall data.

```{r }
sto=10+60+58
print(sto)
print(binom.test(60,sto,0.5))
```

Great.  The probability value is high and the 95% confidence interval for the proportion in the 50% PIs (0.3800 to 0.5589) contains 0.5.  Now let's look for bias.

```{r }
sto=10+58
print(sto)
print(binom.test(10,sto,0.5))
```

Awful! The @RN10592 equation is clearly under-estimating stature.

Now let's use the male "tall" data and the humerus.

```{r fig.dim=c(6,6)}
K_run(i.which="Hum",which.dat=3)
```

Now let's do the binomial test for the proportion in the 50% prediction intervals for the overall data.

```{r }
sto=72+82+17
print(sto)
print(binom.test(82,sto,0.5))
```

Great.  The probability value is high and the 95% confidence interval for the proportion in the 50% PIs (0.4027 to 0.5571) contains 0.5.  Now let's look for bias.

```{r }
sto=72+17
print(sto)
print(binom.test(72,sto,0.5))
```

Awful! The @RN10592 equation is clearly over-estimating stature for the tall (while underestimating stature for the short).  This is the reverse of what we expect for "inverse" calibration (regressing stature on long bone length).  So, apparently the classical calibration has over-corrected.  Let's see how @RN9000 does as that (in theory) will not over-correct.

#### References


